---
title: "Prediction Assignment Writeup"
author: "Ahmed Ameen"
date: "1/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction

This is the model to predict the manner in which subjects did the exercise based on  the data coming from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

## Getting the data

```{r}
set.seed(2021-01-03)
# downloading and saving data
if(!file.exists("./data")){
  dir.create("./data")
}

if(!file.exists("./data/pml-training.csv")){
    fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
    download.file(fileUrl, "./data/pml-training.csv")
}

if(!file.exists("./data/pml-testing.csv")){
    fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
    download.file(fileUrl, "./data/pml-testing.csv")
}
```

## Loadind and  Exploring Data

```{r ,cache=TRUE}

trainData<-read.csv("./data/pml-training.csv")

testData<-read.csv("./data/pml-testing.csv")

# show number of rows and columns 

dim(trainData)

dim(testData)

str(trainData)

str(testData)

```

## Preprocessing Data

As shown above there are columns have many NA values so I need to remove these columns and zero variance and there are columns that are useless in our model (first seven column)

```{r}
library(caret)

# remove columns have average of NA more than 0.5

trainData1 <- trainData[, colMeans(is.na(trainData)) <= .5]

# remove columns have one unique value

nzvTrainData <- nearZeroVar(trainData1)

trainData2 <- trainData1[, -c(1,2,3,4,5,6,nzvTrainData)]

```


After removing columns that have many NAs values and zero variance I need to reduce the number of predictor so I will the correlation between predictors and remove columns that are highly correlated.

```{r}
library(caret)

# get numeric predictors

numericDataTrainData <- trainData2[sapply(trainData2, is.numeric)]

# Remove highly correlated variables (greater than 0.7)

descrCor <- cor(numericDataTrainData)

highlyCorrelatedTrainingData <- findCorrelation(descrCor, cutoff=0.7)

# Get columns name

 highlyCorColTrainData <- colnames(numericDataTrainData)[highlyCorrelatedTrainingData]

 trainData3 <- trainData2[, -which(colnames(trainData2) %in% highlyCorColTrainData)]
 
 names(trainData3)
```

Random forest 

```{r,cache=TRUE}
library(randomForest)

# get numeric predictors

rf <-randomForest(as.factor(classe)~.,data= trainData3, importance=TRUE,ntree=500)
rf

```

Preparing test date to be predicted

```{r,cache=TRUE}
library(dplyr)
# get numeric predictors
 testData1 <- testData %>% rename(classe = problem_id )
testData1 <- testData1[, c(names(trainData3))]
predict(rf, newdata = testData1)
```
